Security vs. Maintainability: Fixing Vulnerabilities Obfuscates your Code

### 2019-01-28 paper/ICPC19/icpc19.tex
s3 finished

\newcommand{\Sof}[1]{\textbf{[Sofia:[}{\color{cyan} #1}\textbf{]]}}

@@

\begin{abstract}
  Security is a crucial non-functionality requirement for software applications.
  However, building secure software is far from trivial as developers lack both
  the knowledge and tools to effectively address this concern. In this paper, we
  study the impact of changes to improve security on the maintainability of several
  open source applications. Using a dataset containing $607$ security-oriented
  commits, we measure maintainability --- as computed by the Software Improvement
  Group's web-based source code analysis service \emph{Better Code Hub} (BCH) ---
  before and after the security refactoring. Results show that making software
  more secure comes at a cost on maintainability. This is particularly evident
  in refactorings to deal with \textit{Broken Authentication} and \textit{Cross-Site Request Forgery} attacks.
  \textcolor{mypink3}{Furthermore, we have found evidence that security-related changes are more
  likely to be modified in the future than regular code changes.}
\end{abstract}

@@

@@

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

@@

Under the rush of companies trying to outdo each other and due to the lack of expertise in
the security field, security flaws are usually detected only when hackers exploit them which happens
because developers create buggy software without knowing. Gladly, fixing these
issues usually is as simple as changing the software codebase. However,
these refactorings can sometimes have a negative impact on software maintenance, not
only because developers are under the rush of "protecting" their software but also because 
they tend to take the easiest path and not the most efficient one. Thus, we
want to understand how security refactorings impact the maintainability of software.

@@

\setcounter{lstannotation}{0}
\begin{lstlisting}[caption={Fix provided by OpenSSL developers to the CVE-2014-3506 vulnerability},label={lst:vuln}]
+static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
+
+ unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
+ if(max_len < (unsigned long)s->max_cert_list) /*!\annotation{lst:func2}!*/
+    return s->max_cert_list;
+ return max_len;
+}

@@

In the \texttt{dtls1\_reassemble\_fragment} function the value of \texttt{msg\_hdr-$>$frag\_off+frag\_len} was checked against the
maximum hadshake message size but \texttt{msg\_len} is allocated to the fragment buffer in line 37. Thus, if the fragment was within
the allowed size, the pending handshake message could consume $16MB$+$2MB$. According with the message commit, 10 handshake messages
are allowed, so the attacker could consume approximately $180MB$ per DTLS connection.

@@

Although limiting the size of a handshake message seems an elementary problem to repair, the example shows that there was a significant amount of changes performed in the codebase which produced a negative impact in the project maintainability. Even though the developer tries to simplify the codebase with the creation of a smaller unit \texttt{dtls1\_max\_handshake\_message\_len}, it also disrupts 2 of the guidelines proposed by the Software Improvement Group (SIG) for building maintainable software~\cite{Visser:2016:OREILLY}. It adds a new branch point to the \texttt{dtls1\_process\_out\_of\_seq\_message} unit (\ref{lst:func5}) increasing the cyclomatic complexity and breaking the \emph{Write Simple Units of Code} guideline. And, it increases the codebase size in 4 LOCs (13 lines added, 9 lines deleted) which infringes the \emph{Keep Your Codebase Small} guideline.

@@

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.5\textwidth]{figures/methodology.pdf}
 	\caption{Study Methodology}
	\label{fig:met}
\end{figure}

@@

% \begin{table}[h]
% 	\centering
% \caption{Current Dataset Patterns Distribution} \label{tab:patterns}
% \begin{tabular}{@{}ll@{}}
% \toprule
% Pattern & \# Commits\\
% \midrule
% Injection & 89\\
% Broken Authentication and Session Management& 45\\
% Cross-Site Scripting& 142\\
% Broken Access Control& 2\\
% Security Misconfiguration& 9\\
% Sensitive Data Exposure& 24\\
% Insufficient Attack Protection& 18\\
% Cross-Site Request Forgery& 34\\
% Using Components With Known Vulnerabilities& 27\\
% Unprotected APIs& 7\\
% Memory Leak& 91\\
% Overflow& 19\\
% Denial-of-Service& 43\\
% Path Traversal& 19\\
% Resource Leak& 33\\
% SHA-1 Hash Function & 1\\
% Miscellaneous& 101\\\midrule
% Total& 716\\
% \bottomrule
% \end{tabular}
% \end{table}

@@

In total, the dataset contains 716 security refactorings. These refactorings were computed by the BCH tool to calculate maintainability. However, some of the refactorings are not considered in the final results due to limitations of BCH: lack of language support and project size. The final dataset contains $607$ security refactorings. The Wilcoxon statistical test - described more in-depth in subsection \ref{sec:statsval} - is performed to evaluate the significance of our results. Each type of security flaw is tested, Figure \ref{fig:pat}. One of the main requirements of this test is that the input sample needs to comprise more than 20 commits. Thus, the patterns not complying with this requirement were reallocated to the \textit{Miscellaneous} pattern. In the end, we highlight the following patterns:

\begin{itemize}
	\item \textbf{Using Components with Known Vulnerabilities.} The majority of software produced today integrates several components, such as libraries and frameworks, which may affect the software if developers use vulnerable versions - usually known and disclosed somewhere on the Internet. \textit{Based on the OWASP Top 10 of 2017.}
	\item \textbf{Broken Authentication \& Session Management.} Uncorrectly implemented functionalities related to authentication and session management, allowing crackers to gain access to session tokens, passwords, keys and other sensitive data \textit{- Based on the OWASP Top 10 of 2013.}
	\item \textbf{Cross-Site Request Forgery.} Poor session tokens generation and management usually allow crackers to send forged HTTP requests including authentication information from the victim to vulnerable web application \textit{- Based on the OWASP Top 10 of 2013.}
	\item \textbf{Denial-of-Service} Security flaws that can allow a cracker to flood or crashing services. This attacks usually occur when the system receives much traffic causing it to slow down and eventually stop. For example, flaws allowing crackers to trigger memory allocations corresponding to large length values (Listing \ref{lst:vuln}).
	\item \textbf{Miscellaneous} This pattern comprises several others categories of security refactorings (e.g., \textit{path traversal and buffer overflow}). It contains several other security refactorings that do not have their own pattern yet. The patterns not satisfying the size requirement of more than 20 refactorings are contemplated here.
	\item \textbf{Cross-Site Scripting.} Lack of proper validation or escaping allow crackers to submit untrusted data to web browsers through malicious scripts that can hijack the user sessions or redirect the user to malicious sites \textit{- Based on the OWASP Top 10 of 2017.}
	\item \textbf{Injection.} When developers do not keep untrusted data separate from commands and queries. If a cracker sends a string that exploits the syntax of the interpreter, then an injection attack is possible (e.g., SQL and LDAP injection)\textit{- Based on the OWASP Top 10 of 2017.}
	\item \textbf{Memory Leak.} Memory management issue found more frequently in programming languages that do not manage memory automatically (e.g., C/C++ and Objective-C), i.e., where instead developers are responsible for handling it. One of the main causes of DoS attacks.
\end{itemize}

@@

Table \ref{tab:dataset} presents the descriptive statistics of the 94 open-source projects involved in this study.

\subsection{Security vs. Baseline Commits}
%
Previous studies tried to measure the impact of refactorings on open-source software maintainability~\cite{HEGEDUS2018313} before. But to the best of our knowledge, none of them used the BCH model and tried to evaluate the impact of security refactorings compared with regular refactorings as we do. Thus, we analyze the maintainability of regular commits and use them as a baseline.

The baseline dataset uses the security commits dataset as input. For each
security commit, one random commit is selected from the list of all commits of
the belonging project. We originate the regular commits from the security
commits to ensure that differences in maintainability are not consequence of
characteristics of different projects.
%
\subsection{Maintainability Analysis}

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.5\textwidth]{figures/guidelines.pdf}
 	\caption{10 guidelines to produce maintainable code}
	\label{fig:guidelines}
\end{figure}

\Sof{=============== READY FINISH!}

The tool calculates the compliance against each particular guideline.

For each guideline, BCH evaluates the compliance against a particular guideline by setting boundaries for the percentage of code allowed to fall in each of the four risk severity categories (low risk, medium risk, high risk, and very high risk). If the thresholds are not violated, the project is considered to be compliant with the guideline. According to BCH, the guideline thresholds are calibrated yearly based on a representative benchmark of closed and open-source software systems. Being compliant with a guideline means that the project under analysis is at least better than 65\% of the software systems in BCH’s benchmark.

The BCH report of the $<INSERT\_EXEMPLO>$ for a non-compliant guideline can be seen in Fig. [X]. This was extracted from the report of the app $<INSERT\_EXEMPLO>$, used in the motivating example of Section II [MAYBE?]. The green bar represents the percentage of compliant lines of code. These lines of code are considered to be compliant with ISO 25010 standard for maintainability [21]. The yellow, orange and red bars represent non-compliant lines of code with medium, high, and very high severity levels, respectively. Along the bars, there are also marks that refer to the compliance thresholds for each severity level. The report is equivalent to the information reported in Table II: a set of thresholds, number of lines of code (LOC), and percentage of the project for each severity level. Nonetheless, thresholds provided by BCH do not sum to 100\%: non-compliant levels are provided in a cumulative way (e.g., the threshold for the medium level includes high and very high levels); the compliant-level threshold is the complement of the medium-level threshold.

Since we want to analyze maintainability regression, we
use BCH to compute maintainability in two different versions of the Android app: a) the version of the project before the security commit ($v_{E-1}$) and b) the version immediately after the security commit ($v_E$). This is illustrated in Fig. [X].
Although BCH provides a detailed report of the maintainability of the project, it does not compute a final score that we can use to compare maintainability amongst different projects. Thus, based in previous work CITE?, we designed an equation to capture the distance between the current state of the project and the standard thresholds. We have adjusted the equation to meet the following requirements:

\begin{itemize}
	\item \textbf{The maintainability difference between two versions of the same project is not affected by its size.} In this work, we want to evaluate the identical security patterns occurring in different projects. Thus, the metric cannot use normalization based on its size – we convert percentage data to the respective number of lines of code.
	\item \textbf{Distance to the thresholds in high severity levels is more penalized than in low severity levels.} We use weights based on the severity level to count lines of code that violate maintainability guidelines.
\end{itemize}

@@

\noindent
where $G$ is the group of selected maintainability guidelines from BCH (e.g., Write short units of code, etc.) and $v$ is the
version of the software under evaluation. The maintenance $M$ based on the guideline $g$ for a given version of a project is computed with the following equation:

@@

\noindent
where $T(l)$ is the threshold in percentage of the lines of code that are accepted to be non-compliant with the guideline for the severity level $l$. This is a standard value defined by BCH. In other words, the factor $w$ is used in Eq. \ref{eq:3} to highlight the lines of code that are not complying with the guideline. For instance, the threshold for the severity level veryHigh is defined in Table II as T(veryHigh) = 6.9\%, which derives to a weight of w(veryHigh) = 13.5. This means that, in this example, one non-compliant guideline is decreasing maintainability score by 13.5 points while a compliant guideline is increasing by 1.0 point. In addition, a version that is perfectly aligned with the standard thresholds has a maintainability score of zero. Then, we compute the difference of maintainability between the security commit ($v_E$) and its parent commit ($v_{E-1}$), as illustrated in Fig. [X].

@@

To validate the maintainability differences in different groups of commits (e.g., baseline and security commits) we use the Paired Wilcoxon signed-rank test with the significance level $\alpha = 0.05$. In other words, we test the null hypothesis that the maintainability difference between pairs of versions $v_{E-1}$, $v_E$ (i.e., before and after an security-commit) follows a symmetric distribution around 0. To understand the effect-size, as advocated by the Common-language effect sizes [X], we compute the mean difference, the median of the difference, and the percentage of cases that reduce maintainability.

\textcolor{mypink3}{@todo: Discuss pratt improvement and accuracy} Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

@@

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

@@

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

@@

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

@@

It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

>>>>

\newcommand{\Sof}[1]{\textbf{[Sofia:[}{\color{red} #1}\textbf{]]}}

@@

\begin{abstract}
  Security is a crucial non-functionality requirement for software applications.
  However, building secure software is far from trivial as developers lack both
  the knowledge and tools to effectively address this concern. In this paper, we
  study the impact of changes to improve security on the maintainability of several
  open source applications. Using a dataset containing 607 security-oriented
  commits, we measure maintainability --- as computed by the Software Improvement
  Group's web-based source code analysis service \emph{Better Code Hub} (BCH) ---
  before and after the security refactoring. Results show that making software
  more secure comes at a cost on maintainability. This is particularly evident
  in refactorings to deal with \textit{Broken Authentication} and \textit{Cross-Site Request Forgery} attacks.
\end{abstract}

@@

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

@@

@@

\Sof{@todo: Improve this paragraph}Under the rush of companies trying to outdo each other and due to the lack of expertise in the security field, security flaws are usually detected only when hackers exploit them which happens
because developers create buggy software without knowing. Gladly, fixing these
issues usually is as simple as changing the software codebase. However,
these refactorings can sometimes have a negative impact on software maintenance, not
only because developers are under the rush of "protecting" their software but also because 
they tend to take the easiest path and not the most efficient one. Thus, we
want to understand how security refactorings impact the maintainability of software.

@@

\setcounter{lstannotation}{0}
\begin{lstlisting}[caption={Fix provided by OpenSSL developers to the CVE-2014-3506 vulnerability \Sof{@todo: Fix the regex issue}},label={lst:vuln}]
+static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
+
+ unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
+ if(max_len < (unsigned long)s->max_cert_list) /*!\annotation{lst:func2}!*/
+    return s->max_cert_list;
+ return max_len;
+}

@@

In the \texttt{dtls1\_reassemble\_fragment} function the value of \texttt{msg\_hdr-$>$frag\_off+frag\_len} was checked against the
maximum handshake message size but \texttt{msg\_len} is allocated to the fragment buffer in line 37. Thus, if the fragment was within
the allowed size, the pending handshake message could consume $16MB$+$2MB$. According with the message commit, 10 handshake messages
are allowed, so the attacker could consume approximately $180MB$ per DTLS connection.

@@

Although limiting the size of a handshake message seems an elementary problem to repair, the example shows that there was a significant amount of changes performed in the codebase which produced a negative impact in the project maintainability. Even though the developer tries to simplify the codebase with the creation of a smaller unit \texttt{dtls1\_max\_handshake\_message\_len}, it also disrupts 2 of the guidelines proposed by the Software Improvement Group (SIG) for building maintainable software~\cite{Visser:2016:OREILLY}. It adds a new branch point to the \texttt{dtls1\_process\_out\_of\_seq\_message} unit (\ref{lst:func5}) increasing the cyclomatic complexity and breaking the \emph{Write Simple Units of Code} guideline. And, it increases the codebase size in $4$ lines of code ($13$ lines added, $9$ lines deleted) which infringes the \emph{Keep Your Codebase Small} guideline.

@@

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.5\textwidth]{figures/methodology.pdf}
 	\caption{Study Methodology \Sof{@todo: Improve it!}}
	\label{fig:met}
\end{figure}

@@

@@

In total, the dataset contains $716$ security refactorings. These refactorings were computed by the BCH tool to calculate maintainability. However, some of the refactorings are not considered in the final results due to limitations of BCH: lack of language support and project size. The final dataset contains $607$ security refactorings. The Wilcoxon statistical test - described more in-depth in subsection \ref{sec:statsval} - is performed to evaluate the significance of our results. Each type of security flaw is tested, Figure \ref{fig:pat}. One of the main requirements of this test is that the input sample needs to comprise more than $20$ commits. Thus, the patterns not complying with this requirement were reallocated to the \textit{Miscellaneous} pattern. In the end, we highlight the following patterns:

\begin{itemize}
	\item \textbf{Using Components with Known Vulnerabilities.} The majority of software produced today integrates several components, such as libraries and frameworks, which may affect the software if developers use vulnerable versions - usually known and disclosed somewhere on the Internet. \textit{Based on the OWASP Top 10 of 2017.}
	\item \textbf{Broken Authentication \& Session Management.} Uncorrectly implemented functionalities related to authentication and session management, allowing crackers to gain access to session tokens, passwords, keys and other sensitive data. \textit{Based on the OWASP Top 10 of 2013.}
	\item \textbf{Cross-Site Request Forgery.} Poor session tokens generation and management usually allow crackers to send forged HTTP requests including authentication information from the victim to vulnerable web application. \textit{Based on the OWASP Top 10 of 2013.}
	\item \textbf{Denial-of-Service} Security flaws that can allow a cracker to flood or crashing services. This attacks usually occur when the system receives much traffic causing it to slow down and eventually stop. For example, flaws allowing crackers to trigger memory allocations corresponding to large length values (Listing \ref{lst:vuln}).
	\item \textbf{Miscellaneous} This pattern comprises several others categories of security refactorings (e.g., \textit{path traversal and buffer overflow}). It contains several other security refactorings that do not have their own pattern yet. The patterns not satisfying the size requirement of more than 20 refactorings are contemplated here.
	\item \textbf{Cross-Site Scripting.} Lack of proper validation or escaping allow crackers to submit untrusted data to web browsers through malicious scripts that can hijack the user sessions or redirect the user to malicious sites. \textit{Based on the OWASP Top 10 of 2017.}
	\item \textbf{Injection.} When developers do not keep untrusted data separate from commands and queries. If a cracker sends a string that exploits the syntax of the interpreter, then an injection attack is possible (e.g., SQL and LDAP injection). \textit{Based on the OWASP Top 10 of 2017.}
	\item \textbf{Memory Leak.} Memory management issue found more frequently in programming languages that do not manage memory automatically (e.g., C/C++ and Objective-C), i.e., where instead developers are responsible for handling it. One of the main causes of DoS attacks.
\end{itemize}

@@

Table \ref{tab:dataset} presents the descriptive statistics of the 94 open-source projects involved in this study. \Sof{@all: Should I expand this? I am not sure what it is important to say about this.}

\subsection{Security vs. Baseline Commits}

Previous studies tried to measure the impact of refactorings on open-source software maintainability~\cite{HEGEDUS2018313} before. But to the best of our knowledge, none of them used the BCH model and tried to evaluate the impact of security refactorings compared with regular refactorings as we do. Thus, we analyze the maintainability of regular commits and use them as a baseline.

The baseline dataset uses the security commits dataset as input. For each
security commit, one random commit is selected from the list of all commits of
the belonging project. We originate the regular commits from the security
commits to ensure that differences in maintainability are not consequence of
characteristics of different projects.

\subsection{Maintainability Analysis}

@@

\begin{figure}[H]
 	\centering 	\includegraphics[width=0.49\textwidth]{figures/guidelines.pdf}
 	\caption{Guidelines to produce maintainable code}
	\label{fig:guidelines}
\end{figure}

All guidelines are measured in the same evaluation by BCH. During each guideline evaluation, the tool determines the compliance towards one guideline by establishing limits for the percentage of code allowed to be in each of the 4 risk severity levels (\emph{low risk}, \emph{medium risk}, \emph{high risk} and \emph{very high risk}). If the project does not violate the thresholds, then it is compliant with the guideline. These thresholds are calibrated by BCH using their own data - open-source and closed software systems. If a project is compliant with a guideline, it means that it is at least $65\%$ better than the software used by BCH to calculate the thresholds\footnote{Check the answer to \emph{How can I adjust the threshold for passing/not passing a guideline?} at https://bettercodehub.com/docs/faq (Accessed on January 27, 2019)}. Figure \ref{fig:bchrep} shows an example of the  guidelines report provided by BCH for a project after finishing its evaluation. These are the results after performing the OpenSSL CVE-2014-3506 vulnerability refactoring - motivation example discussed, previously, in section \ref{sec:motivation}. This version of OpenSSL only complies with 1 out of 10 guidelines, the \emph{Write Clean Code} guideline.

\begin{figure}[H]
 	\centering 	\includegraphics[width=0.45\textwidth]{figures/bch_report.png}
 	\caption{Maintainability report of OpenSSL CVE-2014-3506 vulnerability refactoring for the guideline \emph{Write Simple Units of Code} provided by \emph{Better Code Hub}. This version of OpenSSL does not comply with the guideline in the example since the bars do not reach the threshold points. This example only complies with 1/10 guidelines (\emph{Write Clean Code}).}
	\label{fig:bchrep}
\end{figure}

\emph{Units} are the smallest groups of code that can be maintained and executed independently \cite{Visser:2016:OREILLY}. McCabe \cite{1702388} is a software metric that calculates the number of linearly independent paths of the source code. BCH uses this metric to measure the number of branch points in the \emph{Write Simple Units of Code} guideline. The bar represents the top 30 of units that violates the guideline, sorted by severity, which is indicated by the colors of the checkboxes. The green bar represents the number of compliant branch points per unit (\emph{at most 5}), i.e., the units are compliant with ISO 25010 \cite{iso:2011}. Yellow, orange and red bars represent units that do not comply with medium (\emph{above 5}), high (\emph{above 10}) and very high (\emph{above 25}) severity levels. In the bar, there are marks that pinpoint the compliance thresholds for each severity level.

\begin{figure}[H]
 	\centering 	\includegraphics[width=0.49\textwidth]{figures/commit.pdf}
 	\caption{Maintainability difference for security commits}
	\label{fig:commit}
\end{figure}

Aiming to analyze the impact of security refactorings, we use BCH to compute maintainability of two different versions of the project (Figure \ref{fig:commit}):
\begin{itemize}
	\item $v_{s-1}$, the version containing the security flaw, i.e., before the refactoring;
	\item $v_{s}$, the version after fixing the security flaw, i.e., after the refactoring;
\end{itemize}

The BCH tool does not compute the final score that our study needs to compare maintainability amongst different project versions. Thus, our work proposes an equation to capture the distance between the current state of the project and the standard thresholds based on previous work \cite{Olivari:2018}. 
\begin{itemize}
	\item \textbf{Size project changes do not affect the maintainability difference between two versions, $\Delta M (v_{s-1},v_{s})$.} In this work, we aim to evaluate the alike security patterns occurring in different projects. Thus, normalization based on its size cannot be used - we convert relative data to the respective number of lines of code.
	\item \textbf{Distance to lower severity levels are less penalized than to the thresholds in high severity levels.} Severity level weights based on the severity level to count lines of code that violate maintainability guidelines.
\end{itemize}

@@

\noindent
where $G$ is the group of selected maintainability guidelines from BCH (Figure \ref{fig:guidelines}) and $v$ is the
version of the software under evaluation. The maintenance $M$ based on the guideline $g$ for a given version of a project is computed with the following equation:

@@

\noindent
where $T(l)$ is the threshold in percentage of the lines of code that are accepted to be non-compliant with the guideline for the severity level $l$. This is a standard value defined by BCH. In other words, the factor $w$ is used in Equation \ref{eq:3} to highlight the lines of code that are not complying with the guideline. Then, we compute the difference of maintainability between the security commit ($v_{s}$) and its parent commit ($v_{s-1}$), as illustrated in Figure \ref{fig:commit}.

@@

To validate the maintainability differences in different groups of commits (e.g., baseline and security commits), we use the Paired Wilcoxon signed-rank test \cite{10.2307/3001968} with the significance level $\alpha = 0.05$. In other words, we test the null hypothesis that the maintainability difference between pairs of versions $v_{v-1}$, $v_v$ (i.e., before and after a security-commit) follows a symmetric distribution around $0$. To understand the effect-size, as advocated by the Common-language effect sizes \cite{graw:1992}, we compute the mean difference, the median of the difference, and the percentage of cases that reduce maintainability.

@@

@@

@@

@@
